---
title: "R Notebook"
output: html_notebook
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


```{r}
# CARREGAR O PACOTE
library(data.table)
# LEITURA DA BASE
base =
  fread(
    input = paste0("dados_salarios.csv"),
    header = T,
    na.strings = "NA",
    data.table = FALSE,
    dec = ","
  )

# FORMATACAO
library(dplyr)
base$trab_remoto = recode(base$trab_remoto,
                          `0` = "1:Não",
                          `50` = "2:Parcial",
                          `100` = "3:Total")

base$experiencia = recode(
  base$experiencia,
  `EN` = "1:EN",
  `MI` = "2:MI",
  `SE` = "3:SE",
  `EX` = "4:EX"
)

write.csv2(base, "base.csv", row.names = FALSE)

```

```{r}
# CARREGAR TABELA FORMATADA E SALVA
library(data.table)
base =
  fread(
    input = paste0("base.csv"),
    header = T,
    na.strings = "NA",
    data.table = FALSE,
    dec = ","
  )
```

```{r}
# UMA VARIAVEL QUALITATIVA: TABELA DE FREQUENCIA
freq.tabela = table(base$experiencia, useNA = "ifany")
freq.tabela
porc.tabela = round(prop.table(freq.tabela) * 100, 1)
porc.tabela
freq.tabela = data.frame(freq.tabela, porc.tabela)
freq.tabela = freq.tabela[,-3]
colnames(freq.tabela) = c("Experiencia", "Frequencia", "Porcentagem")
```
```{r}
# GRAFICO DE BARRAS "EXPERIENCIA"
# GRAFICO DE SETORES CIRCULARES "EXPERIENCIA"
png(file = "barras.png")
barplot(
  height = freq.tabela$Frequencia,
  names = freq.tabela$Experiencia,
  col = rgb(0.3, 0.6, 0.5, 0.5),
  xlab = "Experiência",
  ylab = "Frequência",
  main = ""
)
dev.off()

png(file = "setores.png")
pie(freq.tabela$Frequencia, freq.tabela$Experiencia,
    main = "Tempo de Experiência")
dev.off()
```


```{r}
# DUAS VARIAVEIS QUALITATIVAS: TABELA CRUZADA OU GRAFICO DE BARRAS "EXPERIENCIA VS TRAB_REMOTO"
freq.tabela =
  table(base$experiencia, base$trab_remoto, useNA = "ifany")
freq.tabela
porc.tabelaL = round(prop.table(freq.tabela, 1) * 100, 1)
porc.tabelaL
tabela = data.frame(table(base$experiencia, base$trab_remoto))
colnames(tabela) = c("Experiencia", "Trab_Remoto", "Freq")
```

```{r}
# GRAFICOS FREQUENCIA E PORCENTAGEM
library(ggplot2)
ggplot(tabela, aes(fill = Trab_Remoto, y = Freq, x = Experiencia)) +
  geom_bar(position = "dodge", stat = "identity")

ggplot(tabela, aes(fill = Trab_Remoto, y = Freq, x = Experiencia)) +
  geom_bar(position = "fill", stat = "identity") +
  ylab("Porcentagem")
```
```{r}
# UMA VARIAVEL QUANTITATIVA: HISTOGRAMA "SALARIO_USD"
hist(base$salario_USD)
hist(base$salario_USD,
     breaks = 6)
hist(
  base$salario_USD,
  main = "Distribuição Salarial",
  xlab = "Salários",
  col = "darkmagenta",
  freq = FALSE
)
```
```{r}
# UMA VARIAVEL QUALITATIVA E UMA VARIAVEL QUANTITATIVA:
# TABELA COM MEDIAS "EXPERIENCIA VS SALARIO_USD"
tabela.medias =
  aggregate(base$salario_USD,
            by = list(base$experiencia),
            FUN = "mean")
colnames(tabela.medias) = c("Experiencia", "Sal_Medio")

library(ggplot2)
ggplot(tabela.medias, aes(x = Experiencia, y = Sal_Medio)) +
  geom_bar(stat = "identity")
```
```{r}
# DUAS VARIAVEIS QUANTITATIVAS: GRAFICO DE DISPERSAO "BASE NOVA"
ggplot(data = iris, aes(Petal.Length, Petal.Width)) +
  geom_point()
png(file = "dispersao.png")
ggplot(data = iris, aes(Petal.Length, Petal.Width, color = Species)) +
  geom_point()
dev.off()
```
```{r}
# DUAS VARIAVEIS QUANTITATIVAS, SENDO UMA DELAS O TEMPO: GRAFICO DE LINHAS "ANO VS SALARIO_USD"
salario_ano =
  aggregate(base$salario_USD, by = list(base$ano), FUN = "mean")
colnames(salario_ano) = c("ano", "salario")
ggplot(data = salario_ano, aes(x = ano, y = salario)) +
  geom_line(color = "blue", lwd = 1.8) +
  geom_point(color = "red", lwd = 6)
```
```{r}
# TRANSFORMACAO DE UMA VARIAVEL QUANTITATIVA PARA VARIAVEL QUALITATIVA "SALARIO_USD"
library(psych)
base$Cat_Salario[base$salario_USD <100000]  = "G1"
base$Cat_Salario[base$salario_USD >=100000 & base$salario_USD <200000]  = "G2"
base$Cat_Salario[base$salario_USD >=200000 & base$salario_USD <300000]  = "G3"
base$Cat_Salario[base$salario_USD >=300000 & base$salario_USD <400000]  = "G4"
base$Cat_Salario[base$salario_USD >=400000]  = "G5"
freq.tabela = table(base$Cat_Salario, useNA = "ifany") 
freq.tabela

write.csv2(base, "base.csv", row.names = FALSE)
```

